### 2021-12-03 docs/manuscript/manuscript.tex
Include references in the main .tex

\bibliography{referencias-tesis}{}

>>>>

\begin{thebibliography}{10}

\bibitem{Turing1952}
Turing AM.
\newblock {The chemical basis of morphogenesis}.
\newblock Bulletin of Mathematical Biology. 1952;52(1-2):153--197.
\newblock doi:{10.1007/BF02459572}.

\bibitem{Wolpert1969}
Wolpert L.
\newblock {Positional information and the spatial pattern of cellular
  differentiation}.
\newblock Journal of Theoretical Biology. 1969;25(1):1--47.
\newblock doi:{https://doi.org/10.1016/S0022-5193(69)80016-0}.

\bibitem{sharpe_2019}
Sharpe J.
\newblock Wolpert's {French} {Flag}: what's the problem?
\newblock Development (Cambridge, England). 2019;146(24).
\newblock doi:{10.1242/dev.185967}.

\bibitem{driever_gradient_1988}
Driever W, Nüsslein-Volhard C.
\newblock A gradient of bicoid protein in {Drosophila} embryos.
\newblock Cell. 1988;54(1):83--93.
\newblock doi:{10.1016/0092-8674(88)90182-1}.

\bibitem{driever_bicoid_1988}
Driever W, Nüsslein-Volhard C.
\newblock The bicoid protein determines position in the {Drosophila} embryo in
  a concentration-dependent manner.
\newblock Cell. 1988;54(1):95--104.
\newblock doi:{10.1016/0092-8674(88)90183-3}.

\bibitem{affolter_decapentaplegic_2007}
Affolter M, Basler K.
\newblock The {Decapentaplegic} morphogen gradient: from pattern formation to
  growth regulation.
\newblock Nature Reviews Genetics. 2007;8(9):663--674.
\newblock doi:{10.1038/nrg2166}.

\bibitem{dessaud_dynamic_2010}
Dessaud E, Ribes V, Balaskas N, Yang LL, Pierani A, Kicheva A, et~al.
\newblock Dynamic {Assignment} and {Maintenance} of {Positional} {Identity} in
  the {Ventral} {Neural} {Tube} by the {Morphogen} {Sonic} {Hedgehog}.
\newblock PLOS Biology. 2010;8(6):e1000382.
\newblock doi:{10.1371/journal.pbio.1000382}.

\bibitem{cohen_theoretical_2014}
Cohen M, Page KM, Perez-Carrasco R, Barnes CP, Briscoe J.
\newblock A theoretical framework for the regulation of {Shh}
  morphogen-controlled gene expression.
\newblock Development. 2014;141(20):3868--3878.
\newblock doi:{10.1242/dev.112573}.

\bibitem{raspopovic_digit_2014}
Raspopovic J, Marcon L, Russo L, Sharpe J.
\newblock Digit patterning is controlled by a {Bmp}-{Sox9}-{Wnt} {Turing}
  network modulated by morphogen gradients.
\newblock Science. 2014;345(6196):566.
\newblock doi:{10.1126/science.1252960}.

\bibitem{Milo2002}
Milo R, Shen-Orr SS, Itzkovitz S, Kashtan N, Chklovskii D, Alon U.
\newblock {Network Motifs: Simple Building Blocks of Complex Networks}.
\newblock Science. 2002;298(5594):824--827.
\newblock doi:{10.1126/science.298.5594.824}.

\bibitem{kitano_computational_2002}
Kitano H.
\newblock Computational systems biology.
\newblock Nature. 2002;420(6912):206--210.
\newblock doi:{10.1038/nature01254}.

\bibitem{Kalir2005}
Kalir S, Mangan S, Alon U. {A coherent feed-forward loop with a SUM input
  function prolongs flagella expression in Escherichia coli}; 2005.

\bibitem{Kalir2004}
Kalir S, Alon U.
\newblock {Using a quantitative blueprint to reprogram the dynamics of the
  flagella gene network}.
\newblock Cell. 2004;117(6):713--720.
\newblock doi:{10.1016/j.cell.2004.05.010}.

\bibitem{Mangan2003a}
Mangan S, Zaslaver A, Alon U.
\newblock {The coherent feedforward loop serves as a sign-sensitive delay
  element in transcription networks.}
\newblock Journal of molecular biology. 2003;334(2):197--204.

\bibitem{ODonnell2005}
O'Donnell KA, Wentzel EA, Zeller KI, Dang CV, Mendell JT.
\newblock {c-Myc-regulated microRNAs modulate E2F1 expression.}
\newblock Nature. 2005;435(7043):839--843.
\newblock doi:{10.1038/nature03677}.

\bibitem{Basu2005}
Basu S, Gerchman Y, Collins CH, Arnold FH, Weiss R.
\newblock {A synthetic multicellular system for programmed pattern formation}.
\newblock Nature. 2005;434(7037):1130--1134.
\newblock doi:{10.1038/nature03461}.

\bibitem{atkinson_development_2003}
Atkinson MR, Savageau MA, Myers JT, Ninfa AJ.
\newblock Development of genetic circuitry exhibiting toggle switch or
  oscillatory behavior in {Escherichia} coli.
\newblock Cell. 2003;113(5):597--607.
\newblock doi:{10.1016/s0092-8674(03)00346-5}.

\bibitem{friedland_synthetic_2009}
Friedland AE, Lu TK, Wang X, Shi D, Church G, Collins JJ.
\newblock Synthetic gene networks that count.
\newblock Science (New York, NY). 2009;324(5931):1199--1202.
\newblock doi:{10.1126/science.1172005}.

\bibitem{gardner2012}
Gardner L, Deiters A.
\newblock Light-controlled synthetic gene circuits.
\newblock Current Opinion in Chemical Biology. 2012;16(3-4):292--299.
\newblock doi:{10.1016/j.cbpa.2012.04.010}.

\bibitem{levskaya2005}
Levskaya A, Chevalier AA, Tabor JJ, Simpson ZB, Lavery LA, Levy M, et~al.
\newblock Engineering {Escherichia} coli to see light.
\newblock Nature. 2005;438(7067):441--442.
\newblock doi:{10.1038/nature04405}.

\bibitem{santos-moreno_using_2019}
Santos-Moreno J, Schaerli Y.
\newblock Using {Synthetic} {Biology} to {Engineer} {Spatial} {Patterns}.
\newblock Advanced Biosystems. 2019;3(4):1800280.
\newblock doi:{https://doi.org/10.1002/adbi.201800280}.

\bibitem{karlsson_therapeutic_2012}
Karlsson M, Weber W.
\newblock Therapeutic synthetic gene networks.
\newblock Current Opinion in Biotechnology. 2012;23(5):703--711.
\newblock doi:{10.1016/j.copbio.2012.01.003}.

\bibitem{higashikuni_advancing_2017}
Higashikuni Y, Chen WC, Lu TK.
\newblock Advancing therapeutic applications of synthetic gene circuits.
\newblock Current Opinion in Biotechnology. 2017;47:133--141.
\newblock doi:{10.1016/j.copbio.2017.06.011}.

\bibitem{abil_synthetic_2015}
Abil Z, Xiong X, Zhao H.
\newblock Synthetic biology for therapeutic applications.
\newblock Molecular Pharmaceutics. 2015;12(2):322--331.
\newblock doi:{10.1021/mp500392q}.

\bibitem{healy_genetic_2019}
Healy CP, Deans TL.
\newblock Genetic circuits to engineer tissues with alternative functions.
\newblock Journal of Biological Engineering. 2019;13:39.
\newblock doi:{10.1186/s13036-019-0170-7}.

\bibitem{kitada_programming_2018}
Kitada T, DiAndreth B, Teague B, Weiss R.
\newblock Programming gene and engineered-cell therapies with synthetic
  biology.
\newblock Science (New York, NY). 2018;359(6376).
\newblock doi:{10.1126/science.aad1067}.

\bibitem{munteanu_2014}
Munteanu A, Cotterell J, Solé RV, Sharpe J.
\newblock Design principles of stripe-forming motifs: the role of positive
  feedback.
\newblock Scientific Reports. 2014;4(1):5003.
\newblock doi:{10.1038/srep05003}.

\bibitem{Cotterell2010}
Cotterell J, Sharpe J.
\newblock {An atlas of gene regulatory networks reveals multiple three-gene
  mechanisms for interpreting morphogen gradients}.
\newblock Molecular Systems Biology. 2010;6(425):1--14.
\newblock doi:{10.1038/msb.2010.74}.

\bibitem{vonDassow2000}
von Dassow G, Meir E, Munro EM, Odell GM.
\newblock The segment polarity network is a robust developmental module.
\newblock Nature. 2000;406(6792):188--192.
\newblock doi:{10.1038/35018085}.

\bibitem{Reinitz1995}
Reinitz J, Mjolsness E, Sharp DH.
\newblock {Model for cooperative control of positional information in
  Drosophila by bicoid and maternal hunchback}.
\newblock Journal of Experimental Zoology. 1995;271(1):47--56.
\newblock doi:{10.1002/jez.1402710106}.

\bibitem{Jaeger2004}
Jaeger J, Blagov M, Kosman D, Kozlov KN, Manu, Myasnikova E, et~al.
\newblock {Dynamical analysis of regulatory interactions in the gap gene system
  of Drosophila melanogaster}.
\newblock Genetics. 2004;167(4):1721--1737.
\newblock doi:{10.1534/genetics.104.027334}.

\bibitem{Schaerli2018}
Schaerli Y, Jim{\'{e}}nez A, Duarte JM, Mihajlovic L, Renggli J, Isalan M,
  et~al.
\newblock {Synthetic circuits reveal how mechanisms of gene regulatory networks
  constrain evolution}.
\newblock Molecular Systems Biology. 2018;14(9):e8102.
\newblock doi:{10.15252/msb.20178102}.

\bibitem{Elowitz2000}
Elowitz MB, Leibier S.
\newblock {A synthetic oscillatory network of transcriptional regulators}.
\newblock Nature. 2000;403(6767):335--338.
\newblock doi:{10.1038/35002125}.

\bibitem{barbier2020}
Barbier I, Perez-Carrasco R, Schaerli Y.
\newblock Controlling spatiotemporal pattern formation in a concentration
  gradient with a synthetic toggle switch.
\newblock Molecular Systems Biology. 2020;16(6):e9361.
\newblock doi:{10.15252/msb.20199361}.

\bibitem{Schaerli2014}
Schaerli Y, Munteanu A, Gili M, Cotterell J, Sharpe J, Isalan M.
\newblock {A unified design space of synthetic stripe-forming networks}.
\newblock Nature Communications. 2014;5(May):4905.
\newblock doi:{10.1038/ncomms5905}.

\bibitem{balaskas2012}
Balaskas N, Ribeiro A, Panovska J, Dessaud E, Sasai N, Page KM, et~al.
\newblock Gene {Regulatory} {Logic} for {Reading} the {Sonic} {Hedgehog}
  {Signaling} {Gradient} in the {Vertebrate} {Neural} {Tube}.
\newblock Cell. 2012;148(1):273--284.
\newblock doi:{10.1016/j.cell.2011.10.047}.

\bibitem{exelby_2021}
Exelby K, Herrera-Delgado E, Perez LG, Perez-Carrasco R, Sagner A, Metzis V,
  et~al.
\newblock Precision of tissue patterning is controlled by dynamical properties
  of gene regulatory networks.
\newblock Development. 2021;148(dev197566).
\newblock doi:{10.1242/dev.197566}.

\bibitem{ricci2011}
Ricci F, Vallée-Bélisle A, Plaxco KW.
\newblock High-precision, in vitro validation of the sequestration mechanism
  for generating ultrasensitive dose-response curves in regulatory networks.
\newblock PLoS computational biology. 2011;7(10):e1002171.
\newblock doi:{10.1371/journal.pcbi.1002171}.

\bibitem{mendoza2006}
Mendoza L, Xenarios I.
\newblock A method for the generation of standardized qualitative dynamical
  systems of regulatory networks.
\newblock Theoretical Biology \& Medical Modelling. 2006;3:13.
\newblock doi:{10.1186/1742-4682-3-13}.

\bibitem{vu2007}
Vu TT, Vohradsky J.
\newblock Nonlinear differential equation model for quantification of
  transcriptional regulation applied to microarray data of {Saccharomyces}
  cerevisiae.
\newblock Nucleic Acids Research. 2007;35(1):279--287.
\newblock doi:{10.1093/nar/gkl1001}.

\bibitem{zhang2013}
Zhang Q, Bhattacharya S, Andersen ME.
\newblock Ultrasensitive response motifs: basic amplifiers in molecular
  signalling networks.
\newblock Open Biology. 2013;3(4):130031.
\newblock doi:{10.1098/rsob.130031}.

\bibitem{armao2016}
Armao JJ, Lehn JM.
\newblock Nonlinear {Kinetic} {Behavior} in {Constitutional} {Dynamic}
  {Reaction} {Networks}.
\newblock Journal of the American Chemical Society. 2016;138(51):16809--16814.
\newblock doi:{10.1021/jacs.6b11107}.

\bibitem{chen2005}
Chen KC, Wang TY, Tseng HH, Huang CYF, Kao CY.
\newblock A stochastic differential equation model for quantifying
  transcriptional regulatory network in {Saccharomyces} cerevisiae.
\newblock Bioinformatics (Oxford, England). 2005;21(12):2883--2890.
\newblock doi:{10.1093/bioinformatics/bti415}.

\bibitem{dahlquist2015}
Dahlquist KD, Fitzpatrick BG, Camacho ET, Entzminger SD, Wanner NC.
\newblock Parameter {Estimation} for {Gene} {Regulatory} {Networks} from
  {Microarray} {Data}: {Cold} {Shock} {Response} in {Saccharomyces} cerevisiae.
\newblock Bulletin of Mathematical Biology. 2015;77(8):1457--1492.
\newblock doi:{10.1007/s11538-015-0092-6}.

\bibitem{skbio2020}
scikit-bio~development team T. scikit-bio: A Bioinformatics Library for Data
  Scientists, Students, and Developers; 2020.
\newblock Available from: \url{http://scikit-bio.org}.

\bibitem{D.2005}
Bonchev DG, Rouvray DH.
\newblock {Complexity in Chemistry, Biology, and Ecology, ser. Mathematical and
  Computational Chemistry.}
\newblock Springer; 2005.

\bibitem{sklearn}
Pedregosa F, Varoquaux G, Gramfort A, Michel V, Thirion B, Grisel O, et~al.
\newblock Scikit-learn: Machine Learning in {P}ython.
\newblock Journal of Machine Learning Research. 2011;12:2825--2830.

\bibitem{Erdos1959}
Erd{\"{o}}s P, R{\'{e}}nyi a.
\newblock {On random graphs}.
\newblock Publicationes Mathematicae. 1959;6:290--297.
\newblock doi:{10.2307/1999405}.

\bibitem{Watts1998}
Watts DJ, Strogatz SH.
\newblock {Collective dynamics of 'small-world' networks.}
\newblock Nature. 1998;393(6684):440--2.
\newblock doi:{10.1038/30918}.

\bibitem{giererMeinh1972}
Gierer A, Meinhardt H.
\newblock A theory of biological pattern formation.
\newblock Kybernetik. 1972;12(1):30--39.
\newblock doi:{10.1007/BF00289234}.

\bibitem{meinhardt2000}
Meinhardt H, Gierer A.
\newblock Pattern formation by local self-activation and lateral inhibition.
\newblock BioEssays: News and Reviews in Molecular, Cellular and Developmental
  Biology. 2000;22(8):753--760.
\newblock doi:{10.1002/1521-1878(200008)22:8<753::AID-BIES9>3.0.CO;2-Z}.

\bibitem{shoaf1984}
Shoaf SA, Conway K, Hunt RK.
\newblock Application of reaction-diffusion models to cell patterning in
  {Xenopus} retina. {Initiation} of patterns and their biological stability.
\newblock Journal of Theoretical Biology. 1984;109(3):299--329.
\newblock doi:{10.1016/s0022-5193(84)80085-5}.

\bibitem{ishihara_2006}
Ishihara S, Kaneko K.
\newblock Turing pattern with proportion preservation.
\newblock Journal of Theoretical Biology. 2006;238(3):683--693.
\newblock doi:{10.1016/j.jtbi.2005.06.016}.

\bibitem{kitano2004}
Kitano H.
\newblock Biological robustness.
\newblock Nature Reviews Genetics. 2004;5(11):826--837.
\newblock doi:{10.1038/nrg1471}.

\bibitem{lo_robust_2015}
Lo WC, Zhou S, Wan FYM, Lander AD, Nie Q.
\newblock Robust and precise morphogen-mediated patterning: trade-offs,
  constraints and mechanisms.
\newblock Journal of The Royal Society Interface. 2015;12(102):20141041.
\newblock doi:{10.1098/rsif.2014.1041}.

\bibitem{wagner_robustness_2008}
Wagner A.
\newblock Robustness and evolvability: a paradox resolved.
\newblock Proceedings of the Royal Society B: Biological Sciences.
  2008;275(1630):91--100.
\newblock doi:{10.1098/rspb.2007.1137}.

\bibitem{ishihara_cross_2005}
Ishihara S, Fujimoto K, Shibata T.
\newblock Cross talking of network motifs in gene regulation that generates
  temporal pulses and spatial stripes.
\newblock Genes to Cells. 2005;10(11):1025--1038.
\newblock doi:{https://doi.org/10.1111/j.1365-2443.2005.00897.x}.

\bibitem{Li2019}
Li L, Rispoli R, Patient R, Ciau-Uitz A, Porcher C.
\newblock {Etv6 activates vegfa expression through positive and negative
  transcriptional regulatory networks in Xenopus embryos}.
\newblock Nature Communications. 2019;10(1).
\newblock doi:{10.1038/s41467-019-09050-y}.

\end{thebibliography}
